---
title: "Spotify"
author: "User"
date: "2024-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lets begin by loading the required packages
```{r}
pacman::p_load(ggplot2, plotly, readxl, lubridate, janitor, dplyr, tidyverse)
```

## lets load the dataset
```{r}
Spotify= read.csv("Spotify_Youtube.csv")
```

## Lets view the first 5 rows of the dataset
```{r}
head(Spotify, 5)
```

## Data Cleaning
### Lets check the columns with missing values
```{r}
NAs = colSums(is.na(Spotify))
print(NAs)
names(Spotify)[NAs>1]
```

### Lets remove the missing value
```{r}
Spotify = Spotify |> filter(!is.na(Spotify$Stream))
Spotify = Spotify |> filter(!is.na(Spotify$Comments))
Spotify = Spotify |>  filter(!is.na(Spotify$Likes))
Spotify = Spotify |> filter(!is.na(Spotify$Danceability))
```

### lets check the dimension of the remaining dataset
```{r}
dim(Spotify)
```

### Lets get an overview of the dataset
```{r}
str(Spotify)
```

### Lets change the datatypes of some columns
```{r}
Spotify$Album_type = as.factor(Spotify$Album_type)
Spotify$Licensed = as.factor(Spotify$Licensed)
Spotify$official_video = as.factor(Spotify$official_video)
```

### Lets get a structure of the dataset
```{r}
summary(Spotify)
```

## Visualizations
```{r}
colors = c("blue","green","pink", "gray", "black", "purple", "violet", "red","yellow","brown")
```
### Lets get the most streamed music plus the name of the artist
```{r}
### getting the top 10 Artist and Track streamed
Most_Streamed = Spotify |>  
  select(Artist, Stream) |>  
  group_by(Artist) |>
  summarise(Total_streams = round((sum(Stream)/1000000000), 2), .groups="drop") |> 
  arrange(desc(Total_streams)) |> 
  head(10)

### Stacked bar plot
plot_ly(
  data = Most_Streamed,
  x = ~Artist,
  y = ~Total_streams,
  type = 'bar',
  marker = list(color =colors),
  text = ~paste0(Total_streams, " Billion"),
  textposition = "inside"
) |> 
  layout(
    title = "Top 10 Most Streamed Artists",
    xaxis = list(title = "Artist"),
    yaxis = list(title = "Streams(In Billions)", range=c(8, max(Most_Streamed$Total_streams)))
  )
```

### Most Streamed Channel
```{r}
Most_Streamed_Channel = Spotify |> 
  select(Channel, Stream) |> 
  group_by(Channel) |> 
  summarise(Total_Streamers = round(sum(Stream)/1000000000), 2) |> 
  arrange(desc(Total_Streamers)) |> 
  head(10)

#Stacked bar plot
plot_ly(data = Most_Streamed_Channel,
        x= ~ Channel,
        y= ~Total_Streamers,
        type ="bar", 
        marker = list(color =colors),
        text = ~paste0(Total_Streamers, " Billion"),
        textposition = "inside"
) |> 
  layout(
    title = "Top 10 Most Streamed Channels",
    xaxis = list(title = "Channels"),
    yaxis = list(title = "Streams(In Billions)", range= c(13, max(Most_Streamed_Channel$Total_Streamers)))
  )
```

## Most produced Album Type
```{r}
Most_produced_Album_Type = Spotify |> 
  select(Album_type) |> 
  group_by(Album_type) |> 
  summarise(Total_count = n())

plot_ly(data = Most_produced_Album_Type,
        values = ~Total_count,
        labels = ~Album_type,
        type = "pie",
        marker = list(color=colors)) |> 
  layout(
    title="Most Produced Album Type"
  )

```

## Track with the most Views, Likes, Comments and Streams,
```{r}
Best_Track = Spotify |> 
  select(Track, Views, Comments, Stream) |> 
  group_by(Track) |> 
  arrange(desc(Views), desc(Stream), desc(Comments)) |> 
  head(10)

print(Best_Track)

```

## Relationship between Licence, Views, Likes and Streams
```{r}
License_Best = Spotify |> 
  select(Licensed, Views, Likes, Stream) |> 
  group_by(Licensed) |> 
  summarise(Total_Views = round(sum(Views)/1000000,2), Total_Likes = round(sum(Likes)/1000000,2), Total_Streams = round(sum(Stream)/1000000, 2))

License_Best_Long = License_Best |> 
  pivot_longer(cols = c("Total_Views","Total_Likes","Total_Streams"),
               names_to = "Popularity",
               values_to = "Totals")

plot_ly(data = License_Best_Long,
        x= ~ Popularity,
        y= ~ Totals,
        color = ~Licensed,
        type = "bar",
        text = ~paste0(Totals, " Billions"),
        textposition = "inside"
) |> 
  layout(
    title = "Relationship Between License, Likes, Comments and Streams",
    xaxis = list(title = "Popularity"),
    yaxis = list(title = "Total in Billions)")
  )
```

## Relationship between official Video, Views, Likes and Streams
```{r}
official_video_Best = Spotify |> 
  select(official_video, Views, Likes, Stream) |> 
  group_by(official_video) |> 
  summarise(Total_Views = round(sum(Views)/1000000,2), Total_Likes = round(sum(Likes)/1000000,2), Total_Streams = round(sum(Stream)/1000000, 2))

official_Best_Long = official_video_Best |> 
  pivot_longer(cols = c("Total_Views","Total_Likes","Total_Streams"),
               names_to = "Popularity",
               values_to = "Totals")

plot_ly(data = official_Best_Long,
        x= ~ Totals,
        y= ~ Popularity,
        color = ~official_video,
        type = "bar",
        text = ~paste0(Totals, " Billions"),
        textposition = "inside"
) |> 
  layout(
    title = "Relationship Between Official Video, Views, Likes and Streams",
    xaxis = list(title = "Total in Billions"),
    yaxis = list(title = "Popularity")
  )
```

## A Heatmap to check the correlation between Views, Likes, Streams and Comments
```{r}
#Lets extract the columns we need
Spotify_subset <- Spotify |> 
  select(Danceability, Energy, Key, Loudness, Speechiness, Acousticness, Instrumentalness, Liveness, Valence,Tempo, Duration_ms, Views, Likes, Comments, Stream) |> 
  cor()

# Heatmap
plot_ly(
  x = colnames(Spotify_subset), 
  y = rownames(Spotify_subset), 
  z = Spotify_subset, 
  type = "heatmap", 
  colors = c("blue", "white", "red")
) |> 
  layout(
    title = "Correlation Heatmap",
    xaxis = list(title = "Variables"),
    yaxis = list(title = "Variables")
  )
```